BioTime

### 2021-01-15 analysis/01_filter_by_traits.Rmd
Update file saving to outside cache.

```{r}
biotime_data <- pins::pin_get("biotime-data", board = "github")
#load(here::here("data", "biotime_data.rda"))
load(here::here("data", "elton_mamm.rda"))
load(here::here("data", "elton_bird.rda"))
load(here::here("data", "amphibio.rda"))

@@

pins::pin(bt_traitfiltered, board = "github")
```

@@

pins::pin(trait_ref, board = "github")
```

@@

usethis::use_data(study_table)
pins::pin(study_table, board = "github")
```

>>>>

```{r}
pins::board_register_github(name = "github", repo = "karinorman/biodivTS_data", branch = "master")

biotime_data <- pins::pin_get("biotime-data", board = "github")
elton_mamm <- pins::pin_get("elton-mamm", board = "github")
elton_bird <- pins::pin_get("elton-bird", board = "github")
amphibio <- pins::pin_get("amphibio", board = "github")

@@

#save locally and in cache
usethis::use_data(bt_traitfiltered)
pins::pin(bt_traitfiltered, board = "github")
```

@@

#save locally and in cache
usethis::use_data(trait_ref)
pins::pin(trait_ref, board = "github")
```

@@

#save locally and in cache
usethis::use_data(study_table)
pins::pin(study_table, board = "github")
```

### 2021-01-15 data-raw/02_biotime_ids.Rmd
Update file saving to outside cache.

Import occurrence data from Biotime

Access database
```{r}
#personal computer
con <- dbConnect(
  drv = RMariaDB::MariaDB(), 
  dbname = "biotime",
  username = "root",
  password = "Ka1172353!", 
  host = "localhost"
)

#thelio
con <- DBI::dbConnect(RMariaDB::MariaDB(), 
                      host="mariadb", 
                      user="root", 
                      password="password", 
                      dbname="biotime")

biotime_sql <- tbl(con, "allrawdata") %>% 
  collect() %>% 
  rename_all(.funs = tolower) %>%
  select(-id_all_raw_data)
```
```{r}
data <- biotime_sql %>% 
  group_by(id_species, plot, year, study_id) %>% 
  summarise(year_abun = sum(na.omit(abundance)), 
            year_biomass = sum(na.omit(biomass))) %>%
  left_join(tbl(con, "datasets") %>% select(taxa = TAXA, study_id = STUDY_ID) %>% collect()) %>%
  left_join(tbl(con, "species") %>% select(id_species = ID_SPECIES, genus_species = GENUS_SPECIES) %>% collect()) %>%
  ungroup()

#should be zero to be appropriately aggregated
#data %>% group_by(study_id, plot, year, id_species) %>% filter(n()>1) %>% View()
```

Replace NA plots with a plot label (these are studies with no plot subdivision)
```{r}
#replace plot NA's that are really studies with only one plot with a plot label 
data <- data %>% 
  mutate(plot = case_when(
    is.na(plot) ~ "A",
    TRUE ~ plot
  ))
```

Load data from outside processing script
```{r}
load(here::here('data/BioTIME_grid_filtered.Rdata'))
data <- bt_grid_filtered %>%
  rename_with(tolower) %>%
  mutate(species = str_replace(species, "_", " "))

@@

Check which authority would result in the most matches 
```{r}
##Counts show that OTT is the best authority 
get_match_counts(data %>% filter(taxa == "Birds"), "genus_species")
get_match_counts(data, "genus_species")
```

OTT is the best authority but doesn't have common names, so we use the next best ITIS. We'll start with BioTime data:

##Biotime

>>>>

## Get Biotime Occurrence

### Option for Getting data from SQL query (not grid filtered)

Access database
```{r}
# #personal computer
# con <- dbConnect(
#   drv = RMariaDB::MariaDB(), 
#   dbname = "biotime",
#   username = "root",
#   password = "Ka1172353!", 
#   host = "localhost"
# )
# 
# #thelio
# con <- DBI::dbConnect(RMariaDB::MariaDB(), 
#                       host="mariadb", 
#                       user="root", 
#                       password="password", 
#                       dbname="biotime")
# 
# biotime_sql <- tbl(con, "allrawdata") %>% 
#   collect() %>% 
#   rename_all(.funs = tolower) %>%
#   select(-id_all_raw_data)
# ```
# ```{r}
# data <- biotime_sql %>% 
#   group_by(id_species, plot, year, study_id) %>% 
#   summarise(year_abun = sum(na.omit(abundance)), 
#             year_biomass = sum(na.omit(biomass))) %>%
#   left_join(tbl(con, "datasets") %>% select(taxa = TAXA, study_id = STUDY_ID) %>% collect()) %>%
#   left_join(tbl(con, "species") %>% select(id_species = ID_SPECIES, genus_species = GENUS_SPECIES) %>% collect()) %>%
#   ungroup()
# 
# #should be zero to be appropriately aggregated
# #data %>% group_by(study_id, plot, year, id_species) %>% filter(n()>1) %>% View()
```

Replace NA plots with a plot label (these are studies with no plot subdivision)
```{r}
# #replace plot NA's that are really studies with only one plot with a plot label 
# data <- data %>% 
#   mutate(plot = case_when(
#     is.na(plot) ~ "A",
#     TRUE ~ plot
#   ))
```

### Or, from the grid processing script
```{r}
#file created by 01_biotime_grid.R
load(here::here('data/BioTIME_grid_filtered.Rdata'))
data <- bt_grid_filtered %>%
  rename_with(tolower) %>%
  mutate(species = str_replace(species, "_", " "))

@@

### Now get id's
Check which authority would result in the most matches 
```{r}
##Counts show that OTT is the best authority 
get_match_counts(data %>% filter(taxa == "Birds"), "genus_species")
get_match_counts(data, "genus_species")
```

OTT is the best authority but doesn't have common names, so we use the next best ITIS
